---
title: "final_project_full_team"
author: "TEAM Shukla-Kang-Lee-Mittel-Ding"
date: "`r format(Sys.time())`"
output: html_document
---

Code chunk for set up and initial data load/tidying:
```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)
library(plotly)
library(leaflet)

knitr::opts_chunk$set(
echo = TRUE,
warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
```{r}
nypd_shooting_df = 
  read_csv("data/nypd_shooting_data.csv") %>%
  janitor::clean_names() %>%
  separate(col = occur_date, into = c("month", "day", "year"), sep = "/") %>% 
  separate(col = occur_time, into = c("hour", "minute", "second"), sep = ":") %>%
  mutate(across(where(is.character), tolower),
         month = as.numeric(month),
         month_name = recode(month, "1" = "january", "2" = "february", "3" = "march", "4" = "april", "5" = "may", "6" = "june", "7" = "july", "8" = "august", "9" = "september", "10" = "october", "11" = "november", "12" = "december"),
         day = as.numeric(day),
         year = as.numeric(year), 
         hour = as.numeric(hour),
         minute = as.numeric(minute), 
         second = as.numeric(second),
         minute_calc = hour * 60 + minute,
         boro = as.factor(boro), 
         boro = fct_relevel(boro, "manhattan", "brooklyn", "bronx", "queens", "staten island")) %>%
  select(incident_key, year, month_name, month, day, hour, minute, second, minute_calc, everything())
```

**Aaki additions:**
Scatterplot with location of shooting - text label contains month, year, victim’s age group, perp’s age group - spacing is not great, need to fix
```{r}
nypd_shooting_df %>%
  mutate(text_label = str_c("Month", month_name, "Year", year, "Victim's Age Group", vic_age_group, "Perpetrator's Age Group", perp_age_group)) %>% 
  plot_ly(
    x = ~latitude, y = ~longitude, type = "scatter", mode = "markers",
    color = ~year, text = ~text_label, alpha = 0.5)
```

Box plot with shootings by boro by year - would have loved another continuous variable instead of year, but couldn’t identify a good one
```{r}
nypd_shooting_df %>% 
  mutate(boro = fct_reorder(boro, year)) %>% 
  plot_ly(y = ~year, color = ~boro, type = "box", colors = "viridis")
```

Cluster map:
```{r}
nypd_shooting_df %>%
  leaflet() %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(lat = ~latitude, lng = ~longitude, radius = .25) %>%
  addMarkers(
  clusterOptions = markerClusterOptions())
```

**Jen Lee:**
Similar as aboveoverlaid over NYC map using leaflet library (added that to set up code chunk), text label has month, year, boro, location description
```{r}
pal <- colorFactor("viridis", nypd_shooting_df$year)

  nypd_shooting_df %>%
  mutate(
    text_label = str_c(month_name, " ", year, ", ", boro, ", ", location_desc)) %>%
  leaflet() %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(lat = ~latitude, lng = ~longitude, radius = .1, color = ~pal(year), label = ~text_label) %>%
  addLegend("bottomright", pal = pal, values = ~year,
    title = "year")
```

Stacked bar plot showing similar information as above, shootings by boro by year
```{r}
nypd_shooting_df %>%
  group_by(year, boro) %>%
  summarise(
    count = n()) %>%
  ggplot(aes(fill = boro, y = count, x = year)) +
  geom_bar(position = "stack", stat = "identity") + 
  labs(title = "Number of shootings by boro from 2006-2021")
```

Line plot showing similar information as above, shootings by boro by year
```{r}
nypd_shooting_df %>%
group_by(year, boro) %>%
  summarise(
    count = n()) %>%
  ggplot(aes(y = count, x = year, group = boro, color = boro)) +
  geom_point() + geom_line() +
  facet_grid(cols = vars(boro)) + 
  labs(title = "Number of shootings by boro from 2006-2021")
```

Bar plot with number of shootings by boro - if this is right, BK is more dangerous than I thought!
```{r}
nypd_shooting_df %>% 
  count(boro) %>% 
  mutate(boro = fct_reorder(boro, n)) %>% 
  plot_ly(x = ~boro, y = ~n, color = ~boro, type = "bar", colors = "viridis")
```

**Elise Kang:**
Line plot showing similar information, shootings by season (summer = highest, winter = lowest)
```{r}
nypd_shooting_seasons = nypd_shooting_df %>%
  mutate(
    season = case_when(
      month %in% 9:11 ~ "Fall",
      month %in% c(12, 1, 2) ~ "Winter",
      month %in% 3:5 ~ "Spring",
      TRUE ~ "Summer")) 

nypd_shooting_seasons %>%
  group_by(year, season) %>%
  summarise(
    count = n()) %>%
ggplot(aes(y = count, x = year, group = season, color = season)) +
  geom_point() + geom_line() +
  facet_grid(cols = vars(season)) + 
  labs(title = "Number of shootings by season from 2006-2021")
```